Lawvere-Tierney Sheafification\\ in Homotopy Type Theory

### 2015-01-13 paper/sheaf_lics.tex

@@

We now need to show that $\square$ is universal. The following
lemma is central in the proof~:
\begin{lem}
  Let $T:\Type_{n+1}$. Then $\square T$ is the colimit of the closed
  diagonal diagram
  $$\xymatrix{ 
    \cdots \overline{\Delta_3} \tar[r] & \overline{\Delta_2} \dar[r] &
    \overline{\Delta_1}
  }$$
where $\Delta_k$ is the $k$-pullback of $\id : T \to T$.
\end{lem}

@@

Now that $\square T$ is a colimit, we can easily define an inverse to
$\lambda\,(f:\square T \to U),~x \circ \mu_T$ for any separated type
$U$, using the universal property of the colimit.

@@

Let $f,g$ be two maps from $\sum_{e:E} (\chi\,e)$ to $\sum_{x:A}
(Bx)$, equal when precomposed with $\pi_1$.

As $A$ is separated, for any $x:E$, the first components of $f x$ and
$g x$ are equal. For the second components, notice that for all $x:E$,
$\sum_{y:E} x = y$ has a dense subobject, $\sum_{y:\sum_{e:E} (\chi\,
  e)} x=y_1$. 
Using the separation property with these type, one can show the second
components of $f x$ and $g x$ are equals too, with respect to the
first path.

>>>>

We have to be careful~: if we want to propagate the properties of the
first modality $\modal_{-1}$, all the higher modalities $\modal_n$ must be compatible
with $\modal_{-1}$, \ie we want the property
\begin{prop}
  If $P:\HProp$, and $\widehat P$ is $P$ seen as a $\Type_n$, then $\modal_n \widehat P = \modal_{-1} P$, and the
  following coherence diagram commute 
  $$\xymatrix{
    \modal_{-1} P \ar@{=}[r] & \modal_n \widehat P \\
    P \ar@{=}[r] \ar[u]^{\eta_{-1}} & \widehat P \ar[u]_{\eta_n}
  }$$
\end{prop}

@@

We now need to show that $\square$ is universal. The following
lemma is central in the proof~:
\begin{lem}\label{lem:sepiscolim}
  Let $T:\Type_{n+1}$. Then $\square T$ is the colimit of the closed
  diagonal diagram
  $$\xymatrix{ 
    \cdots \overline{\Delta_3} \tar[r] & \overline{\Delta_2} \dar[r] &
    \overline{\Delta_1}
  }$$
where $\Delta_k$ is the $k$-pullback of $\id : T \to T$.
\end{lem}

@@

Now, let $Q$ be a separated $\Type_{n+1}$, and $f:T \to Q$. Then the
following diagram commutes
$$\xymatrix{ 
    \cdots \overline{\Delta_3} \tar[r] \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
    \dar[r] \ar[d]^{f\circ \pi_1} &
    \overline{\Delta_1} = T \ar[ld]^{f\circ \pi_1}\\
    & Q &
  }$$
But we know (lemma~\ref{lem:sepiscolim}) that $\square T$ is the
colimit of the closed diagonals diagram, thus there is an universal
arrow $\square T \to Q$.

This proves the following proposition
\begin{prop}
  $(\square,\mu)$ defines a reflective subuniverse on $\Type_{n+1}$.
\end{prop}

@@

Let $f,g$ be two maps from $\sum_{e:E} (\chi\,e)$ to $\sum_{x:A}
(Bx)$, equal when precomposed with $\pi_1$.
$$\xymatrix{
  \sum_{e:E} (\chi\, e) \dar[r]^{f\circ\pi_1}_{g\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
  E \dar[ru]^f_g&
}$$
We can restrict the previous diagram to 
$$\xymatrix{
  \sum_{e:E} (\chi\, e) \dar[rr]^{\pi_1\circ f\circ\pi_1}_{\pi\circ g\circ \pi_1} \ar[d]_{\mathrm{dense}}&& A \\
  E \dar[rru]^{\pi_1\circ f}_{\pi_1\circ g}&
},$$ 
and as $A$ is separated, $\pi_1\circ f = \pi_2 \circ g$.
For the second components, let $x:E$. Notice that for all $x:E$,
$\sum_{y:E} x = y$ has a dense subobject, $\sum_{y:\sum_{e:E} (\chi\,
  e)} x=y_1$~:

$$\xymatrix{
  \sum_{y:\sum_{e:E} (\chi\,
  e)} x=y_1 \dar[rrr]^{\pi_2\circ f\circ\pi_1\circ \pi_1}_{\pi_2\circ g\circ \pi_1\circ \pi_1} \ar[d]_{\mathrm{dense}}&&& B\,x \\
  \sum_{y:E} x = y \dar[rrru]^{\pi_2\circ f\circ \pi_1}_{\pi_2\circ g\circ \pi_1}&
}$$
Using the sepatation property of $B\,x$, one can show that second
components, transported correctly along the first components equality,
are equal. The complete proof can be found in the formalisation.
This proves the following proposition
\begin{prop}
  $(\square,\mu)$ defines a modality on $\Type_{n+1}$.
\end{prop}

As this modality is just a step in the construction, we will not show
that it is left exact, we will do it only for the sheafification modality.
